% ************************************************************
% RESULTS: FIRST-PASS RAPID INDEX
% ************************************************************
\section{Results}

The first-pass extraction identified several public indicators suitable for rapid WASH--health risk stratification. Across assessed displacement sites, environmental-health hazards were widely reported, including visible rodents or pests, sewage in surrounding streets, accumulated solid waste, and flooding or stagnant water. Reported skin infections, rashes, scabies, lice, bedbugs, and related ectoparasitic infestations provided an additional site-level health signal. These indicators were treated as public-health risk signals rather than individual-level disease incidence estimates.

The initial priority index classified the Gaza Strip aggregate, Khan Younis, and assessed displacement sites as high-priority strata. The Gaza Strip aggregate received the maximum score across all five domains because public reporting documented concurrent WASH service stress, environmental-health hazards, constrained health-system capacity, displacement vulnerability, and operational constraints affecting service continuity. Khan Younis also received a maximum score, driven by area-specific reporting of sewage flooding and untreated wastewater accumulation, alongside Gaza-wide WASH and operational constraints. Assessed displacement sites received a high-priority score, with conservative scoring for health-system access and operational constraints pending direct site-level linkage to facility functionality data.

Sensitivity analysis did not change the priority classification. All three strata remained high priority under equal-weight, WASH-weighted, and health-access-weighted specifications. This stability suggests that the first-pass classification is not solely an artifact of one weighting choice. The result should be interpreted as a rapid prioritization signal for follow-up data validation and geospatial refinement, not as causal evidence of morbidity or mortality burden.
% ************************************************************
% END RESULTS: FIRST-PASS RAPID INDEX
% ************************************************************
