% ************************************************************
% RESULTS
% ************************************************************

\section{Results}

\subsection{Evidence-control posture}

The current evidence layer identifies one core household validator, two local hotspot validators, two indicative trend validators, one infrastructure and recovery validator, one background mechanism validator, and one restricted source pending geography resolution. This posture supports manuscript development but does not yet authorize page-specific quantitative claims because exact page binding remains pending for the admissible claims.

\subsection{Allowed claim families}

The core household validator may support assessed-frame statements about household WASH access, sanitation conditions, water availability, and service-condition stress. Local enumeration-area factsheets may support place-specific illustrations only when the relevant geography is named. Light-touch monitoring reports may support time-stamped directional trend language only. The infrastructure and recovery validator may support damage, disruption, and recovery-context statements, but not household liters-per-person-per-day outcome claims. Background health-mechanism literature may support plausible environmental-health pathways, but not current operational prevalence claims.

\subsection{Blocked or restricted material}

The Shelter-WASH-FSS IDP assessment is currently blocked for Gaza findings until project geography is resolved. It should not appear in the manuscript results section unless the source audit confirms Gaza-specific content and updates the source-claim matrix.

\subsection{Current manuscript finding status}

At this execution stage, the principal result is not a final epidemiologic estimate. The principal result is a validated source-control structure that prevents scope drift. Findings will be promoted into page-specific manuscript results only after source files are attached or official public URLs are recorded and exact claim pages are entered in the audit matrix.

% ************************************************************
% END RESULTS
% ************************************************************
