% ************************************************************
% DISCUSSION
% ************************************************************

\section{Discussion}

The current evidence base supports a diagnostic interpretation rather than a definitive household-prevalence assessment. Bound sources allow the manuscript to compare national and pre-crisis WASH baselines with localized water-quality evidence, political-economy context, health-care-facility baseline conditions, and official time-stamped situation reporting. The same source-control layer prevents overextension into current Gaza-wide household estimates because the joint household assessment, local factsheets, original light-touch monitoring reports, RDNA table source, and restricted IDP source remain unbound.

JMP-derived baseline evidence indicates that country-level WASH indicators were high before the current field-assessment gaps. The occupied Palestinian territory 2024 baseline includes high estimates for at least basic drinking water, at least basic sanitation, and basic hygiene, while safely managed water and sanitation estimates are lower. These values provide a baseline reference point, not a current Gaza household-access estimate.

The pre-crisis inequality evidence sharpens the diagnostic interpretation. The 2020 JMP inequality extract shows substantial Gaza-region dependence on tanker-truck drinking water and very limited piped-into-dwelling drinking-water use. These patterns support a vulnerability interpretation: service stress is likely to be amplified where households depend on delivered or non-networked water arrangements. The inequality evidence remains a baseline frame and should not be read as a current displacement-site measure.

Localized water-quality evidence supports a public-health mechanism rather than a Gaza-wide prevalence estimate. Turkman et al. provide sampled evidence from Khan Younis and the Middle Area during May--June 2024. Within that scope, the source supports discussion of microbial contamination, water-quality degradation, and plausible waterborne-disease pathways. It cannot validate conditions outside the sampled frame.

The political-economy evidence indicates that WASH system performance cannot be interpreted only as an engineering problem. Fayad et al. support a sector-level interpretation in which governance, financing, access, institutional fragmentation, and service-delivery constraints shape system performance. This context strengthens the manuscript argument that technical repair, while necessary, is not sufficient as a full explanatory model.

Health-care-facility baseline data add a service-continuity dimension. The JMP WinHCF extract provides country-level 2023 estimates for facility water, hygiene, waste-management, and environmental-cleaning services. These data support discussion of infection-prevention and service-continuity vulnerabilities, while remaining outside the category of Gaza-specific crisis-damage measurement.

Official OCHA situation reports provide time-stamped context and should be used only within that function. They may anchor the manuscript's description of the reporting environment and broad WASH-system stress, but they do not replace the unbound original monitoring rounds or RDNA tables.

The principal limitation is evidentiary incompleteness at the household and local-factsheet level. Claims C001--C006 and C008 remain blocked. Therefore, the manuscript should not report current Gaza household WASH prevalence, local Al Mawasi or As Sawarah case findings, original LTM trends, RDNA table findings, or IDP-specific findings until those source rows are bound and page locations are recorded.

Taken together, the bound evidence supports a baseline-to-crisis diagnostic manuscript. The strongest defensible contribution is not a final epidemiologic estimate, but a source-gated diagnostic model that separates baseline conditions, localized mechanisms, political-economy constraints, health-care-facility context, official situation reporting, and unresolved evidence gaps.

% ************************************************************
% END DISCUSSION
% ************************************************************
