% ************************************************************
% DISCUSSION
% ************************************************************
\section{Discussion}

The first-pass index classified all analytic strata as high priority, reflecting convergence across WASH service stress, environmental-health hazards, health-system constraints, population vulnerability, and operational constraints. The stability of classifications across weighting specifications suggests that the result is not dependent on a single scoring assumption. The most defensible interpretation is that the available public data identify clear areas for follow-up validation and recovery-prioritization analysis.

This framework is aligned with health-equity and health-systems research because it focuses on variation in exposure, service access, and recovery need. Its practical value lies in converting fragmented public reporting into an auditable prioritization structure that can be updated as more precise geospatial, facility-functionality, and WASH dashboard data become available.

The method is transferable to other conflict-affected or humanitarian WASH settings when four minimum data elements are available: a time-stamped public source, a defined geography, at least one WASH or environmental-health indicator, and a documented rule for converting source evidence into a domain score. The index does not require complete clinical surveillance to be useful as an early prioritization tool. It does, however, require clear separation between observed service or exposure signals and claims about individual-level health outcomes. In other settings, the domain structure can be retained while indicators are adapted to local reporting systems, facility data, displacement-site assessments, or WASH coordination dashboards.
% ************************************************************
% END DISCUSSION
% ************************************************************
